\Large\textbf{Advection versus diffusion in brain ventricular transport}

### 2025-04-02 paper/main.tex
Minor edits in abstract

\begin{abstract}
Cerebrospinal fluid (CSF) is integral to brain function, as the fluid
provides mechanical support for the brain and the flow of CSF helps
distribute nutrients, neurotransmitters and metabolites throughout
the central nervous system. CSF flow is partly driven
by beating of motile cilia, hair-like structures that line the ependymal cell layer
of the walls of the brain ventricles. Despite the physiological importance of CSF flow,
the underlying mechanisms of cilia-mediated flow and transport
in the brain ventricles remain yet to be comprehensively resolved.
In this study, the role of motile cilia in brain intraventricular CSF flow and transport
is analyzed and evaluated. The geometry of the zebrafish larval brain ventricles,
for which detailed knowledge of cilia properties and CSF motion velocities
and frequencies exists, was used to develop finite element methods that model flow and transport.
The computational model is validated by in vivo experiments that monitor transport
of a photoconvertible protein secreted in the brain ventricles.
The results show that while cilia contribute to advection of large particles,
diffusion plays a significant role in the transport of small molecules such as proteins.
It is also demonstrated that cilia location and the geometry of the ventricular system impact solute distribution.
Altogether, this work presents a computational framework that can be applied to other
ventricular systems, together with new concepts of how molecules are transported within the brain and its ventricles.

>>>>

\begin{abstract}
Cerebrospinal fluid (CSF) is integral to brain function, as the fluid
provides mechanical support for the brain and the flow of CSF helps
distribute nutrients, neurotransmitters and metabolites throughout the
central nervous system. CSF flow is partly driven by beating of motile
cilia, hair-like structures that line the ependymal cell layer of the
walls of the brain ventricles. Despite the physiological importance of
CSF flow, the underlying mechanisms of cilia-mediated flow and
transport in the brain ventricles remain yet to be comprehensively
resolved. This study analyzes and evaluates the role of motile cilia
in brain intraventricular CSF flow and transport. The geometry of the
zebrafish larval brain ventricles, for which detailed knowledge of
cilia properties and CSF motion velocities and frequencies exists, was
used to develop finite element methods for modelling flow and
transport. The computational model is validated by in-vivo experiments
that monitor transport of a photoconvertible protein secreted in the
brain ventricles. Our results show that while cilia contribute to
advection of large particles, diffusion plays a significant role in
the transport of small molecules \mer{such as proteins}. We also
demonstrate how cilia location and the geometry of the ventricular
system impact solute distribution. Altogether, this work presents a
computational framework that can be applied to other ventricular
systems, together with new concepts of how molecules are transported
within the brain and its ventricles.
